---
title: "Data analysis"
output: html_notebook
---

```{r echo=FALSE}
library(tidyverse)
library(lmerTest)
import::from(broom.mixed, tidy)
import::from(multcomp, glht, mcp)
import::from(ggbeeswarm, geom_beeswarm)
import::from(modelr, add_residuals, add_predictions)
import::from(qqplotr, stat_qq_line, stat_qq_point, stat_qq_band)
import::from(nortest, lillie.test)
import::from(ARTool, art, artlm, artlm.con, art.con)
source("../util/plot_glht.R")
```

# Preprocess input data

## Load data
```{r}
path_background <- "data/participant_info.csv"
path_log_root <- "data/log"

df_background <- read_csv(path_background, col_types = "cl")

filenames <- list.files(path_log_root, pattern = "*.csv")

df_info <- 
  tibble(
  file_path = list.files(path_log_root, pattern = "*.csv", full.names = TRUE)) %>% 
  mutate(
    filenames = basename(file_path)) %>% 
  separate(filenames, c("user_id", "start_timestamp", "condition"), sep = "_") %>% 
  mutate(
    user_id = str_replace(user_id, "User", ""),
    condition = str_replace(condition, ".csv", ""),
    start_timestamp = as.double(str_replace(start_timestamp, "T", ""))) 

df_log <- 
  df_info %>% 
  mutate(log_df = map(file_path, ~read_csv(.x, col_types = "icdicc"))) %>% 
  unnest(log_df) %>% 
  select(-file_path, start_timestamp)
```

## Calculate WPM
```{r}
df_wpm <- 
  df_log %>% 
  mutate(words = (str_length(transcribed_string) - 1.0) / 5.0) %>% 
  select(user_id, condition, stimulus_index, input_time_ms, words) %>% 
  group_by(user_id, condition, stimulus_index) %>% 
  summarize(
    minute = (last(input_time_ms) - first(input_time_ms)) / 1000.0 / 60.0,
    wpm = first(words) / minute,
    .groups = "drop")
```

## Join with the background data and assign factor data columns
```{r}
df_wpm_bkgnd <- 
  df_wpm %>% 
  left_join(df_background, by = "user_id") %>% 
  mutate(
    condition = factor(condition, levels = c("Sentences", "Random", "Mix")),
    typing_style = factor(if_else(is_touchtypist, "Touchtypist", "Non-touchtypist"), 
                          levels = c("Touchtypist", "Non-touchtypist"))) %>% 
  select(user_id, typing_style, condition, stimulus_index, wpm)

sample_n(df_wpm_bkgnd, 20)
```


# Visualize data

```{r}
df_wpm_bkgnd %>% 
  ggplot(aes(x = condition, y = wpm)) +
  geom_beeswarm(alpha = 0.5) +
  facet_wrap(~typing_style)
```

The data may not be Gaussian-distributed.

```{r}
df_wpm_bkgnd %>% 
  ggplot(aes(x = condition, y = wpm, color = typing_style)) +
  geom_violin() +
  facet_wrap(~user_id)
```


# Analysis without transformation

## Fit model
```{r}
m_ris <- lmer(wpm ~ typing_style + condition + (1 + condition | user_id), data = df_wpm_bkgnd)
m_ris
```

## ANOVA
```{r}
anova(m_ris)
```

## GLHT
```{r fig.height=2, fig.width=6}
m_ris %>% 
  glht(mcp(condition = "Tukey",
           typing_style = "Tukey")) %>% 
  confint() %>% 
  tidy() %>% 
  plot_glht(unit = "WPM", adjustment_method = "Tukey HSD, family-wise")
```
# Assumption check

```{r}
plot(m_ris)

df_aug <- 
  df_wpm_bkgnd %>% 
  add_residuals(m_ris) %>% 
  add_predictions(m_ris)

df_aug %>% 
  ggplot(aes(sample = resid)) +
  stat_qq_band() + 
  stat_qq_line() + 
  stat_qq_point(size = 0.5)

df_aug %>% 
  ggplot(aes(x = pred, y = resid, color = user_id)) +
  geom_hline(yintercept = 0, color = "grey") +
  geom_point(alpha = 0.5) + 
  geom_smooth(aes(group = NULL, color = NULL), method = "loess") +
  facet_grid(condition ~ .) 

```
## Log transformation

```{r}
m_log_ris <- lmer(log(wpm) ~ typing_style + condition + (1 + condition | user_id), data = df_wpm_bkgnd)
anova(m_log_ris)
m_log_ris %>% 
  glht(mcp(condition = "Tukey",
           typing_style = "Tukey")) %>% 
  confint() %>% 
  tidy() %>% 
  mutate(
    estimate = exp(estimate),
    conf.low = exp(conf.low),
    conf.high = exp(conf.high)) %>% 
  plot_glht(unit = "times of the baseline WPM", adjustment_method = "Tukey HSD, family-wise", reference_value = 1)
```

```{r}
df_aug_log <- 
  df_wpm_bkgnd %>% 
  add_residuals(m_log_ris) %>% 
  add_predictions(m_log_ris)

df_aug_log %>% 
  ggplot(aes(sample = resid)) +
  stat_qq_band() + 
  stat_qq_line() + 
  stat_qq_point(size = 0.5)

df_aug_log %>% 
  ggplot(aes(x = pred, y = resid)) +
  geom_point(alpha = 0.5) +
  geom_hline(yintercept = 0, color = "black")


lillie.test(df_aug_log$resid)
```

# Alrigned rank transformation
(NOTE: I disabled these chunks because they take long to run. Remove `eval=FALSE, include=FALSE` to run them)

```{r eval=FALSE, include=FALSE}
art_result <- art(wpm ~ typing_style * condition + (1 | user_id), data = df_wpm_bkgnd)

# anova(art_result)   # NOTE: Takes long time to run

# Analysis of Variance of Aligned Rank Transformed Data
# 
# Table Type: Analysis of Deviance Table (Type III Wald F tests with Kenward-Roger df)
# Model: Mixed Effects (lmer)
# Response: art(wpm)
# 
#                                   F Df Df.res  Pr(>F)
# 1 typing_style              0.15532  1     28 0.69648
# 2 condition              2051.08055  2   4459 < 2e-16 ***
# 3 typing_style:condition    1.48392  2   4459 0.22686
# ---
# Signif. codes:   0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

#================================================================================
# Comparison between Kenward-Roger and Satterthwaite
##### Results of Satterthwaite has a weird DenDF
# Type III Analysis of Variance Table with Satterthwaite's method
#                         Sum Sq Mean Sq NumDF  DenDF F value Pr(>F)
# condition               971847  485923     2 32.875  0.5077 0.6065
# typing_style            148649  148649     1  1.750  0.1553 0.7363
# condition:typing_style 1109597  554798     2  8.559  0.5797 0.5806

###### Results of the Kenward-Roger's method has the same degrees of freedom as car::Anova
# Type III Analysis of Variance Table with Kenward-Roger's method
#                         Sum Sq Mean Sq NumDF DenDF F value Pr(>F)
# condition               971847  485923     2  4459  0.5077 0.6019
# typing_style            148649  148649     1    28  0.1553 0.6965
# condition:typing_style 1109597  554798     2  4459  0.5797 0.5601


art_anova_all <- function(art_obj, type = "III", method = c("Satterthwaite", "Kenward-Roger")) {
  tibble(term = colnames(art_obj$aligned.ranks)) %>% 
    mutate(map_dfr(term, function(term) {
      artlm(art_obj, term) %>% 
        as_lmerModLmerTest() %>% 
        anova(type = type, ddf = method[1]) %>%
        as_tibble(rownames = ".term") %>%
        filter(.term %in% c(term, str_replace(term, ":", "")))    # in older version of some unidentified package, the interaction term had colon removed.
    })) %>% 
    select(-.term)
}

art_anova_all(art_result)



art.con(art_result, ~typing_style)

m_lart_result_cond <- artlm.con(art_result, "condition")
plot(m_lart_result_cond)
```



# Manually extract ART result to fit models
```{r eval=FALSE, include=FALSE}
art_result <- art(wpm ~ typing_style * condition + (1 | user_id), data = df_wpm_bkgnd)

# check sum of aligned responses
assertthat::are_equal(sum(art_result$aligned$typing_style), 0)
assertthat::are_equal(sum(art_result$aligned$condition), 0)
assertthat::are_equal(sum(art_result$aligned$`typing_style:condition`), 0)

# check aligned (unranked) with F tests
df_wpm_aligned <- 
  df_wpm_bkgnd %>% 
  mutate(
    wpm_R_typing_style = art_result$aligned$typing_style,
    wpm_R_condition = art_result$aligned$condition,
    wpm_R_interaction = art_result$aligned$`typing_style:condition`)

anova(lmer(wpm_R_typing_style ~ typing_style * condition + (1 | user_id), data = df_wpm_aligned))
anova(lmer(wpm_R_condition ~ typing_style * condition + (1 | user_id), data = df_wpm_aligned))
anova(lmer(wpm_R_interaction ~ typing_style * condition + (1 | user_id), data = df_wpm_aligned))

# augment ART columns and fit models
df_wpart_result <- 
  df_wpm_bkgnd %>% 
  mutate(
    wpm_AR_typing_style = art_result$aligned.ranks$typing_style,
    wpm_AR_condition = art_result$aligned.ranks$condition,
    wpm_AR_interaction = art_result$aligned.ranks$`typing_style:condition`)

art_result_typing_style <-  lmer(wpm_AR_typing_style ~ typing_style * condition + (1 | user_id), data = df_wpart_result)

art_result_condition <-  lmer(wpm_AR_condition ~ typing_style * condition + (1 | user_id), data = df_wpart_result)

art_result_interaction <-  lmer(wpm_AR_interaction ~ typing_style * condition + (1 | user_id), data = df_wpart_result)

anova(art_result_typing_style, ddf= "Kenward-Roger")

```

