---
title: "Frequentist - Likert-style question"
output: html_notebook
---

```{r}
library(MASS)
library(ordinal)
library(emmeans)
library(tidyverse)
```


Dataset: Experiment 1 of "Blinded with Science" study  (between-subjects, 2 conditions)

```{r}
source("dataset_blinded_with_science.R")
dataset <- dataset_blinded_with_science()
```
The order of the difference is controlled by the order of the factor level of the `condition` column in the dataset. By default, the order is alphabetical. In this dataset, all statistical estimates below are in the followin form:

> [graph] - [no graph]

This order happens to make sense. Between the two the no-graph condition could be considered as the baseline.

# Data overview

```{r}
ftable(xtabs(~ effectiveness + condition, data = dataset))
```

```{r}  
dataset %>% 
  # mutate(effectiveness = as.character(effectiveness)) %>% 
  mutate(effectiveness = fct_rev(factor(effectiveness, levels = 1:9))) %>%
  
  # stacked bar plot
  ggplot(aes(x = condition, fill = effectiveness)) +
  geom_bar(position = "stack", stat="count") +
  
  # grey color scale is robust for colorblind
  scale_fill_brewer(palette="Greys", drop = FALSE) +
  
  # horizontal plot
  coord_flip() +
  
  # legend
  guides(fill = guide_legend(reverse = TRUE)) +
  theme(legend.position = "bottom")
```



# 1. Textbook non-parametric Wilcoxon rank-sum test
```{r}
result_wilcox <- 
  wilcox.test(
    effectiveness ~ condition, 
    data = dataset, 
    paired = FALSE,
    conf.int = TRUE)
result_wilcox
```

# 2. A controversial use of a parametric t-test with ordinal data 
```{r}
result_ttest <- 
  t.test(
    effectiveness ~ condition, 
    data = dataset, 
    paired = FALSE,
    conf.int = TRUE)
result_ttest
```


# 3. Ordinal logistic regression
```{r}
model_olr <- 
  dataset %>% 
  mutate(effectiveness = ordered(effectiveness, levels = as.character(1:9))) %>% 
  clm(effectiveness ~ condition,
      data = .)
```

```{r}
anova(model_olr, type = "3")
```

```{r}
emmeans(model_olr,  ~ condition)
```

```{r}
confint(model_olr)
```


