---
title: "Prepare"
output: html_document
date: "2023-04-09"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Install packages that you don't have following RStudio prompts

Look out for yellow banner above the source code!

```{r setup, include=FALSE}
# session 3
library(tidyverse)
library(tidymodels)
library(modelr)
library(brms)
library(tidybayes)
library(ggdist)
library(distributional)
library(ordinal)
library(ggrepel)
library(glue)
library(emmeans)

# + session 2
library(gtools)
library(patchwork)
library(broom)
library(broom.mixed)
library(bayesplot)
library(knitr)
library(rmarkdown)

# + session 1 and 4
library(multcomp)
library(afex)
library(papaja)
```

If RStudio doesn't help you install, try `pacman`

```{r}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(
  tidyverse, 
  tidymodels,
  modelr,
  brms,
  tidybayes,
  ggdist,
  distributional,
  ordinal,
  ggrepel,
  glue,
  emmeans,
  gtools,
  patchwork,
  broom,
  broom.mixed,
  bayesplot,
  knitr,
  rmarkdown,
  multcomp,
  afex,
  papaja
)
```

To trouble-shoot, run `library()` one at a time

Install missing packages manually with `install.packages("the_missing_package")`

You may need to install the packages below for `brms`

```{r}
# install.packages("igraph")
# install.packages("shiny")
```


Run a simple Bayesian model to verify `brms` installation
From: https://cran.r-project.org/web/packages/brms/readme/README.html

```{r}
fit1 <- brm(count ~ zAge + zBase * Trt + (1|patient),
            data = epilepsy, family = poisson())
summary(fit1)
```

