---
title: "Session 3 Uncertainty visualizations"
author: "Xiaoying Pu, Lace Padilla"
date: "3/3/2022"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(tidymodels)
library(modelr)
library(brms)
library(tidybayes)
library(ggdist)
library(distributional)
library(ordinal)
library(ggrepel)
library(glue)
library(dabestr)
library(emmeans)

theme_set(theme_tidybayes())
```

## Introduction

We will revisit Datasets 1 from last time, focusing on exploring how to
make **visualizations** instead of doing the analysis. By the end of
this tutorial, you will learn to:

1.  Reason about the **expressiveness** and **usability** of plots you
    make.
2.  Identify the **types of uncertainty** from model outputs.
3.  Use resources to **visualize uncertainty** in different ways.

### Expectations

I will run R code and use a Miro board in this tutorial. You can access
the same code in your Rstudio Cloud. Ideally, you will follow along and
run code with me. There are many activities where you can
**participate** and ask **questions** (any time!).

**TIP**: RStudio shows document outline


Miro board (but don't go there yet): <https://miro.com/app/board/uXjVO8L3bt8=/>

------------------------------------------------------------------------

## Activity 1: exploring data

First, we will evaluate some plots made for Dataset 1.

### Load the data

We will base our example on the `blinded.csv` dataset. It's from an
experiment about whether showing people graphs increase belief in drug
efficacy, presumably because graphs are associated with science.

**TIP**: `()` enclosing an assignment will print the assigned value as
well.

```{r 1.0}
# Note that there are 123 rows!

(df1 <- read_csv("data/blinded.csv", show_col_types = FALSE) %>%
  filter(experiment == 1) %>%
  mutate(
    condition = factor(condition), 
    participant_id = paste0("P", str_pad(row_number(), 2, pad = "0"))) %>%
   select(-experiment)
)
```

### Evaluate three plots

Last time, we modeled how `condition` (graph or no graph) affected
`effectiveness`, a Likert scale rating from 1 to 9. We are going to
evaluate three plots:

**Plot 1**

```{r 1.1}
df1 %>%
  ggplot(aes(
    x = effectiveness, 
    y = condition)) +  
  geom_point() +
  scale_x_continuous(breaks = 1:9)
```

**REFLECTION**: is this plot good? Rate (in Zoom chat) its
expressiveness and usability in the Zoom chat on a scale of 1-5

-   *Expressiveness*: the **plot** should represent all relationships
    that exist in the **data**, and those relationships only.
-   *Usability*: the **user** should be able to make accurate inferences
    from the data.

👉ACTION: Enter in Zoom chat: "E3, U3"

**Plot 2**

```{r 1.20}
jitterer <- position_jitter(height = 0.1, seed = 1)
```

```{r 1.21}
df1 %>%
  ggplot(aes(
    x = effectiveness, 
    y = condition,
  )) +     
  geom_point(position = jitterer) +       
  scale_x_continuous(breaks = 1:9)
```

**Reflection**: 👉 how would you rate the expressiveness and usability?

**Concept**: *intrinsic uncertainty* in the data (sans the artificial
jitter!)

**Plot 3**

👉 How else would you plot this data to increase
expressiveness/usability?

One option:

```{r 1.30}
(df1_prop <- df1 %>%
  count(condition, effectiveness) %>%
  add_row(condition = "graph", effectiveness = 2, n = 0) %>%
  complete(condition, effectiveness, fill = list(n = 0)) %>%
  group_by(condition) %>%
  mutate(prob = n/sum(n))
)
```

```{r 1.31}
df1_prop %>%
  ggplot(aes(x = effectiveness, y = n, color = condition)) + 
  geom_point() + 
  geom_line() +
  scale_x_continuous(breaks = 1:9) + 
  scale_y_continuous("Count") + 
  NULL
```

**Reflection**: why is this better? why is it worse?



## Activity 2: visualizing model outputs

### Load and summarize frequentist and Bayesian models

```{r 2.0}
# model objects
m_t_test <- lm(effectiveness ~ - 1 + condition, data = df1)
m_t_test_bayesian <- readRDS("../02-bayesian_stats/rds/dataset1.brm.bayesiant.rds")

# means
means_t_freq <- m_t_test %>% 
   tidy() %>%
   mutate(condition = str_remove(term, "condition"), .before = 1) %>%
   select(-term)

means_t_bayes <- data_grid(df1, condition) %>%
  tidybayes::add_epred_draws(m_t_test_bayesian, newdata = . ,dpar = TRUE)

# preds
preds_t_freq_obj <- predict(m_t_test, newdata = data_grid(df1, condition), 
                     se.fit=TRUE, interval = "prediction", level = 0.95)

preds_t_freq <- m_t_test %>%
  broom::augment(newdata = data_grid(df1, condition), se_fit = TRUE) %>%
  mutate(
    .se.pred = sqrt(.se.fit ^ 2 + preds_t_freq_obj$residual.scale^2), # NOTE
    df = preds_t_freq_obj$df) 

preds_t_bayes <- data_grid(df1, condition) %>%
    tidybayes::add_predicted_draws(m_t_test_bayesian)

```

### Activity 2.1 CI vs. PI

Starting from the point estimate

```{r 2.10}
means_t_freq %>%
  ggplot(aes(x = estimate, y = condition)) + 
  geom_point(aes(effectiveness), position = jitterer, data=df1, alpha = 0.5, color = "#8c96c6") +
  geom_point(position = position_nudge(y = 0.2)) +
  scale_x_continuous("effectiveness", breaks = 1:9, limits = c(0, 10)) + 
  labs(title = "Frequentist lm(), estimated effectiveness by condition") +
  NULL
```

👉ACTION: **head to Miro**
<https://miro.com/app/board/uXjVO8L3bt8=/?moveToWidget=3458764523389765219&cot=14>


**Concepts**

We consider two types of intervals: CI and PI

Frequentist:

-   CI: confidence interval; the long-run proportion of correspondingly
    computed intervals that end up containing the true value of the
    parameter

    -   For our t-test, the 95% CI is from the confidence distribution
        (t-distribution). [More
        reading](https://mjskay.github.io/ggdist/articles/freq-uncertainty-vis.html)
        and an [animated explanation](https://seeing-theory.brown.edu/frequentist-inference/index.html#section2)

-   PI: prediction interval; estimate of an interval in which a future
    observation will fall, with a certain probability, given what has
    already been observed

    -   For our t-test, the PI is also from a t-distribution

Bayesian:

-   CI: credible interval; an interval within which an unobserved
    parameter value falls with a particular probability.
-   PI: posterior predictive interval... from *posterior predictive
    distribution*, the distribution of possible unobserved values
    conditional on the observed values
-   There are no analytical distributions for CIs and PIs here because
    our Bayesian model based on sampling.

Note:

-   CI: parameter uncertainty
-   PI: prediction uncertainty

👉ACTION: **Vote**: is CI always narrower than PI? \[Y/N\]


```{r 2.11}
preds_t_freq %>%
  
  ggplot(aes(y = condition)) + 
  
  # PI
  stat_interval(aes(
    xdist = dist_student_t(
      df = df,
      mu = .fitted, 
      sigma = .se.pred) 
  )) + 
 
  # CI
  stat_pointinterval(aes(
    xdist = dist_student_t(
      df = df,
      mu = .fitted,
      sigma = .se.fit)),
    position = position_nudge(y = -.2)) +
  
  # EXERCISE: put raw data in this plot
  geom_point(
    aes(x = effectiveness),
    data = df1, position = jitterer) +
  
  scale_x_continuous("effectiveness", breaks = 1:9, limits = c(0, 10)) + 
  scale_color_brewer("Predictive interval") + 
  labs(title = "Frequentist lm(), estimated effectiveness by condition",
       subtitle = "Confidence interval widths: 0.66 and 0.95") +
  theme(legend.position="none")
```

👉 Back to Miro and compare our "error bar" guesses with CI/PI we just plotted.

### Activity 2.2 Types of uncertainty vis

👉 Go to Miro
<https://miro.com/app/board/uXjVO8L3bt8=/?moveToWidget=3458764523387189623&cot=14>

### Activity 2.3 Putting it together

A plot showing the confidence distribution for the frequentist t-test. I will 
demonstrate uncommenting code chunks to build up the plot.

**TIP** adding annotation (direct labeling) can increase usability

```{r 2.31}
means_t_freq %>%
  ggplot(aes(y = condition)) +
  
  # raw data
  geom_point(
    aes(
      x = effectiveness,
      y = condition),
    data = df1, position = jitterer, alpha = .5) +
  
  # EXERCISE 1: add an half-eye plot for the confidence distribution
  # stat_halfeye(
  #   aes(xdist =dist_student_t(
  #     df = df.residual(m_t_test),
  #     mu = estimate,
  #     sigma = std.error)),
  #   scale = 0.5, fill = "#8c96c6",
  #   position = position_nudge(y = 0.15)) +
  
  # EXERCIESE 2: change `stat_halfeye` above to `stat_dots`
  
  # EXERCISE 3: add a label for the confidence distribution
  # geom_label_repel(
  #   aes(estimate + 0.25, y = condition),
  #   data = means_t_freq %>% slice_tail(),
  #   label = "CI of the mean", box.padding = 1,
  #   position = position_nudge(y = 0.3), max.overlaps = Inf, seed = 15) +
  
  # EXERCISE 4: add a label for the raw data
  # geom_label_repel(
  #   aes(x = estimate, y = condition),
  #   data = means_t_freq %>% slice_head(),
  #   label = "Jittered raw data", box.padding = 1,
  #   position = position_nudge_repel(y = -0.2), seed = 1) +
  
  scale_x_continuous(breaks = 0:10, limits = c(0, 10)) + 
  labs(title = "Frequentist lm(), confidence distribution of effectiveness by condition") + 
  NULL

```

Recap the plot above:

-   Intrinsic uncertainty: the spread of the raw data (ignore the
    jitter!)
-   Parameter uncertainty: confidence distribution (density plot) and CI
    (point interval)

For the frequentist t-test, showing all types of uncertainty with
different uncertainty vis types!

👉 ACTION: uncomment the chunks below one at a time

```{r 2.32}
preds_t_freq %>%
  ggplot(aes(y = condition)) + 
  
  # Raw data
  # geom_point(
  #   aes(
  #     x = effectiveness,
  #     y = condition),
  #   data = df1, position = jitterer, shape = 21, size =2) +
  
  # Conf. distribution
  # stat_slab(
  #   aes(xdist = dist_student_t(
  #     df = df, 
  #     mu = .fitted, 
  #     sigma = .se.pred)),
  #   slab_color="#9ECAE1", fill = NA, scale = 0.7) +
  
  # Pred. distribution
  # stat_pointinterval(            
  #   aes(xdist =dist_student_t(
  #     df = df.residual(m_t_test), 
  #     mu = estimate, 
  #     sigma = std.error)),
  #   data = means_t_freq,
  #   scale = 0.5, fill = "#8c96c6",
  #   position = position_nudge(y = 0.15)) + 
  
  scale_x_continuous(breaks = 1:9) +
  labs(title = "Frequentist lm(), effectiveness by condition",
       subtitle = "Curves show the prediction distributions\nIntervals show confidence intervals") + 
  NULL
```

👉 ACTION: answer the questions below:

-   What's the intrinsic uncertainty in data?
-   What's been shown with the density curve?
-   What's been shown with the point interval?

This type of plot is sometimes called Kruschke style in **Bayesian**
context
<https://solomonkurz.netlify.app/post/2018-12-20-make-rotated-gaussians-kruschke-style/>.

```{r 2.33}
means_t_bayes %>%
  
  # Sample draws because there are thousands
  
  sample_draws(30) %>%
  
  ggplot(
    aes(y = condition)) + 
 
  # Raw data 
  geom_point(
    aes(x = effectiveness), data = df1,
    shape = 21, size = 2,
    position = jitterer) +
  
  # distributions implied by posterior predictive draws
  stat_slab(
    aes(xdist = dist_student_t(nu, mu, sigma)),
    slab_color="#9ECAE1", alpha = 0.1, fill = NA, scale = 0.7) +
  
  scale_x_continuous(breaks = 1:9, limit = c(0, 11)) + 
  labs(title = "Bayesian t-test, effectiveness by condition",
       subtitle = "Curves show both PI and CI") + 
  NULL
```

**REFLECTION**: 

- Why does the Kruschke spaghetti density show both
parameter and prediction uncertainty? 
- How else can you show PI for the Bayesian model?

-------

**Recap**: 

- Should you show parameter or prediction uncertainty,
or both? 
- Should you use intervals or distributions for uncertainty?


Further reading: 

- CI vs. PI: Hofman, J. M., Goldstein, D. G., & Hullman, J. (2020,
April). How visualizing inferential uncertainty can mislead readers
about treatment effects in scientific results. In Proceedings of the
2020 chi conference on human factors in computing systems (pp. 1-12).
- Interval vs. distribution: Helske, J., Helske, S., Cooper, M., Ynnerman, A., 
& Besançon, L. (2020). Are You Sure You're Sure?-Effects of Visual 
Representation on the Cliff Effect in Statistical Inference. CoRR.




**EXERCISE**: can you make the plot above to look like Activity 2.1?

```{r 2.34}
# Your code here
```




## Packages Information

```{r}
sessionInfo()
```


