---
title: "chi22-course"
author: "Abhraneel Sarma"
date: "3/3/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(tibble)
library(purrr)
library(tidyr)
library(forcats)
library(ggthemes)


library(broom)
library(broom.mixed)

library(modelr)

library(brms)
library(tidybayes)

library(rethinking)

library(ggdist)
library(ggplot2)

theme_set(theme_ggdist())
```


## Introduction

In this document, we will outline the Bayesian analogs of the statistical analyses described [here](https://github.com/chatchavan/CHI-Course-Transparent-Quant/blob/master/frequentist.Rmd) (and in the previous course lecture).

## Data

Load the data and helper functions.

```{r load-data-and-helper}

# Fumemng -> Abhraneel I'm using the Rproj in the root folder, although my is identical...

dataset = readr::read_csv("02-bayesian_stats/data/blinded.csv")
# dataset = readr::read_csv("~/Documents/Github/chi22-course/data/blinded.csv")

# this function extracts results from different models and generate results of the same format to be used in visualizations
tidy.wrapper = function(model) {
  if (class(model) == "lm") {
    tidy(result_ttest, conf.int = TRUE) %>%
      select(-c(statistic, p.value)) %>%
      mutate(model = "Frequentist") %>%
      select(model, everything())
  } else if (class(model) == "brmsfit") {
    tidy(result_ttest.bayesian) %>%
      filter(effect == "fixed") %>%
      select(c(term:conf.high)) %>%
      mutate(model = "Bayesian") %>%
      select(model, everything())
  } else {
    stop("unknown model class")
  }
}
```

TODO: explain the dataset here

### Let's first look at the data in a table

```{r glance-dataset}

    dataset

```  


```{r plot-dataset, fig.height = 3, fig.width = 7}
dataset %>% 
  mutate(effectiveness = fct_rev(factor(effectiveness, levels = 1:9))) %>%
  
  # stacked bar plot
  ggplot(aes(x = condition, fill = effectiveness)) +
  geom_bar(position = "stack", stat="count") +
  
  # plot data for different experiments as small multiples
  facet_wrap( ~ experiment) +
  
  # grey color scale is robust for colorblind
  scale_fill_brewer(palette="Greys", drop = FALSE) +
  
  # horizontal plot
  coord_flip() +
  
  # legend
  guides(fill = guide_legend(reverse = TRUE)) 
```

As we can see above, the original dataset contains results from four different experiments. For the purposes of this lecture, we will confine ourselves to the first experiment.

```{r filter-data}
exp1.data = dataset %>%
  filter(experiment == 1)

exp1.data
```


## Model 1. Wilcoxon signed rank test

This is a non-parametric test which we will skip for now. Although, there exists Bayesian non-parametric methods, they are more advanced for this lecture

## Model 2. t-test

```{r bayesian-t-test}
# this is the linear model equivalent for the paired sample t-test
result_ttest =
  lm(
    effectiveness ~ condition - 1, 
    data = exp1.data
  )

result_ttest.bayesian = 
  brm(
    effectiveness ~ condition - 1,
    prior = c(
      prior(normal(0, 1), class = "b"), # there's a lot of data so even fairly "strong" priors are going to not matter so much here
      prior(student_t(3, 0, 1), class = "sigma")
    ),
    family = student(),
    seed = 99,
    backend = "cmdstanr", #Fumeng: I have to use this as my backend, 'rstan' backend doesn't work on my laptops...
    data = exp1.data,
    chains = 3, cores = 3
  )
```

Next, we visualise the results, along with a side-by-side comparison of the frequentist estimates:

```{r t-test-compare, echo = FALSE, warning = FALSE}
bind_rows(
  tidy.wrapper(result_ttest),
  tidy.wrapper(result_ttest.bayesian)
) %>%
  ggplot() +
  geom_pointrange(
    aes(
      x = term, 
      y = estimate, 
      ymin = conf.low, 
      ymax = conf.high, 
      color = model
    ), position = position_dodge(width = 0.2)) +
    scale_color_brewer(palette = "Set2")
```







## Model 3. Ordinal Logistic Regression












???? 

```{r}
data(Trolley, package = "rethinking")
Trolley
```



```{r}
# figure from: Statistical rethinking with brms, ggplot2, and the tidyverse: Second edition by A Solomon Kurz
exp1.data %>%
  count(effectiveness) %>%
  mutate(pr_k = n / nrow(exp1.data),
         cum_pr_k = cumsum(pr_k)) %>% 
  ggplot(aes(x = effectiveness, y = cum_pr_k, 
             fill = effectiveness)) +
  geom_line(color = canva_pal("Green fields")(4)[2]) +
  geom_point(shape = 21, color = "grey92", 
             size = 2.5, stroke = 1) +
  scale_fill_gradient(low = canva_pal("Green fields")(4)[4],
                      high = canva_pal("Green fields")(4)[1]) +
  scale_x_continuous(breaks = 1:7) +
  scale_y_continuous("cumulative proportion", 
                     breaks = c(0, .5, 1), limits = c(0, 1)) +
  theme(axis.ticks = element_blank(),
        legend.position = "none")
```

What is the model trying to estimate:

```{r}
# primary data
exp1.data_plot = exp1.data %>%
  count(effectiveness) %>%
  mutate(pr_k = n / nrow(exp1.data)) %>%
  add_row(effectiveness = 2, n = 0, pr_k = 0) %>%
  arrange(effectiveness) %>%
  mutate(cum_pr_k = cumsum(n / nrow(exp1.data))) %>% 
  mutate(discrete_probability = ifelse(effectiveness == 1, cum_pr_k, cum_pr_k - pr_k))

text = exp1.data_plot %>%
  mutate(
    text = effectiveness,
    effectiveness = effectiveness + 0.25,
    cum_pr_k = ifelse(cum_pr_k - 0.05 < 0.065, 0.05, cum_pr_k - 0.05)
  )

exp1.data_plot %>% 
  ggplot(aes(x = effectiveness, y = cum_pr_k)) +
  geom_line(aes(color = cum_pr_k), color = canva_pal("Green fields")(4)[1]) +
  geom_point(shape = 21, colour = "grey92", 
             size = 2.5, stroke = 1) +
  geom_linerange(aes(ymin = 0, ymax = cum_pr_k),
                 alpha = 1/2, color = canva_pal("Green fields")(4)[1]) +
  geom_linerange(aes(x = effectiveness + .025,
                     ymin = ifelse(effectiveness == 1, 0, discrete_probability), 
                     ymax = cum_pr_k),
                 color = "black") +
  # number annotation
  geom_text(data = text, 
            aes(label = text),
            size = 4) +
  scale_color_gradient(low = canva_pal("Green fields")(4)[4],
                       high = canva_pal("Green fields")(4)[1]) +
  scale_x_continuous(breaks = 1:7) +
  scale_y_continuous("cumulative proportion", breaks = c(0, .5, 1), limits = c(0, 1)) +
  theme(axis.ticks = element_blank(),
        axis.title.y = element_text(angle = 90),
        legend.position = "none")
```


```{r}
d_plot %>% 
  ggplot(aes(x = response, y = cum_pr_k,
             color = cum_pr_k, fill = cum_pr_k)) +
  geom_line(color = canva_pal("Green fields")(4)[1]) +
  geom_point(shape = 21, colour = "grey92", 
             size = 2.5, stroke = 1) +
  geom_linerange(aes(ymin = 0, ymax = cum_pr_k),
                 alpha = 1/2, color = canva_pal("Green fields")(4)[1]) +
  geom_linerange(aes(x = response + .025,
                     ymin = ifelse(response == 1, 0, discrete_probability), 
                     ymax = cum_pr_k),
                 color = "black") +
  # number annotation
  geom_text(data = text, 
            aes(label = text),
            size = 4) +
  scale_fill_gradient(low = canva_pal("Green fields")(4)[4],
                      high = canva_pal("Green fields")(4)[1]) +
  scale_color_gradient(low = canva_pal("Green fields")(4)[4],
                       high = canva_pal("Green fields")(4)[1]) +
  scale_x_continuous(breaks = 1:7) +
  scale_y_continuous("cumulative proportion", breaks = c(0, .5, 1), limits = c(0, 1)) +
  theme(axis.ticks = element_blank(),
        axis.title.y = element_text(angle = 90),
        legend.position = "none")
```


```{r, eval = FALSE}
result_olr.bayesian = brm(
  bf(effectiveness ~ condition), 
  prior = c(
    prior(normal(0, 1), class = "Intercept"),
    prior(normal(0, 0.5), class = "b")
  ),
  family = cumulative(),
  data = exp1.data)
```


```{r, fig.height = 2, fig.width = 7}
data_grid(exp1.data, condition) %>%
  add_epred_draws(result_olr.bayesian, ndraws = 100) %>%
  group_by(.draw, condition) %>%
  mutate(cum_pr_k = cumsum(.epred)) %>%
  ggplot() +
  geom_line(aes(x = condition, y = cum_pr_k, group = interaction(.draw, .category)), alpha = 0.1) +
  facet_grid(. ~ .category)
```


```{r}
result_olr <- 
  exp1.data %>% 
  mutate(effectiveness = ordered(effectiveness, levels = as.character(1:9))) %>% 
  ordinal::clm(effectiveness ~ condition,
      data = .)

anova(result_olr, type = "3")

emmeans::emmeans(result_olr,  ~ condition)
confint(result_olr)
```

