---
title: "chi22-course"
author: "Abhraneel Sarma"
date: "3/3/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tibble)
library(purrr)
library(tidyr)
library(brms)
```

## Data

Load the data

```{r}
df = readr::read_csv("~/Documents/Github/chi22-course/data/exp2.csv") %>%
  mutate(condition = condition == 'expansive') %>%
  group_by(participant)

df
```

## Model 1

The first model is the BEST test model as described by Kruschke in the paper *Bayesian estimateion supersedes the t-test*. In this model, $\beta$ indicates the mean difference in the outcome variable between the two groups (in this case, the percent change in the BART scores). We fit different priors on $\beta$ and set different weights on these priors to obtain our posterior estimate.

$$
\begin{align}
y_{i} &\sim \mathrm{T}(\nu, \mu, \sigma) \\
\mu &= \alpha_{0} + \beta * x_i \\
\sigma &= \sigma_{a} + \sigma_{b}*x_i \\
\beta &\sim \mathrm{N}(\mu_{0}, \sigma_{0}) \\
\sigma_a, \sigma_b &\sim \mathrm{Cauchy}(0, 2) \\
\nu &\sim \mathrm{exp}(30)
\end{align}
$$


```{r, eval = FALSE}
fit.brm = brm(
  bf(change ~ condition, sigma ~ condition), 
  prior = c(
    prior(normal(0, 2), class = "b"),
    prior(student_t(3, 0, 5), class = "Intercept"),
    prior(student_t(3, 0, 2), class = "b", dpar = "sigma"),
    prior(student_t(3, 0, 2), class = "Intercept", dpar = "sigma")
  ), 
  chains = 4, cores = 4,
  data = df)

summary(fit.brm)
```



```{r}
sim_trial = function(y) {
  # let y be the actual number of trials that a participant will pump until
  # p be the point of explosion for any given trial
  p = sample(1:128, 1)
  ifelse(p < y, p, y)
}

sim_exp = function(i, N = 10) {
  trials = 1:N
  map_dbl(trials, ~ sim_trial(i))
}

sim_ = function(x, iter = 1e3) {
  # here x is the averahe number of pumps by a participant
  # we can perform a naive grid search approach
  # let y >= x
  for (i in ceiling(x):128) {
    pumps = map(
      1:iter, 
      ~ mean(sim_exp(i)) # simulates one experiment and calculates the average number of pumps for that experiment
    ) # repeats the simulation many times
    if ((mean(map_dbl(pumps, mean))) > x) {
      return(i)
    }
  }
}
```


```{r, eval = FALSE}
df.gen = df %>%
  mutate(
    P10 = map(adjP10, sim_),
    P20 = map(adjP20, sim_),
    P30 = map(adjP30, sim_)
  )

saveRDS(df.gen, "~/Documents/Github/chi22-course/data/exp2-gen.rds")
```


```{r}
df.gen = readRDS("~/Documents/Github/chi22-course/data/exp2-gen.rds") %>%
  mutate(
    P10 = map(P10, ~ unlist(ifelse(is.numeric(.), P10, 128))),
    P20 = map(P20, ~ unlist(ifelse(is.numeric(.), P20, 128))),
    P30 = map(P30, ~ unlist(ifelse(is.numeric(.), P30, 128)))
  ) %>%
  mutate(
    trial_P10 = map(P10, ~ (sim_exp(.))),
    trial_P20 = map(P20, ~ (sim_exp(.))),
    trial_P30 = map(P30, ~ (sim_exp(.)))
  ) %>%
  pivot_longer(
    cols = starts_with("trial"),
    names_to = "trial",
    names_prefix = "trial_P"
  ) %>%
  unnest(c(value)) %>%
  group_by(participant) %>%
  mutate(trial = row_number()) %>%
  select(-c(starts_with("adjP"), "P10", "P20", "P30", "change", "orig"))
```

# Model 2

```{r}
# poisson analysis
fit.poiss = brm(
  bf(value ~ condition), 
  prior = c(prior(normal(0, 1), class = "b"), prior(student_t(3, 3.5, 1), class = "Intercept")), 
  data = df.gen, family = poisson(), 
  chains = 4, cores = 4)
```





