---
title: "intro-BDA-example"
author: "Abhraneel Sarma"
date: '2022-03-23'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggdist)
library(cowplot)

theme_set(theme_ggdist() + 
          theme(
            strip.background = element_blank(),
            plot.title = element_text(vjust = .5),
            axis.line.y = element_blank(),
            axis.title.y = element_blank(),
            axis.ticks.y = element_blank(),
            legend.position = 'none'
          ))
```

Recall that the experiment shows participants the effectiveness of a certain drug, which is presented through either text ("No Graph" condition) or through a bar chart ("Graph" condition). Participants are asked to rate the effectiveness of a drug using a 9-item likert scale.

```{r}
dataset = readr::read_csv("02-bayesian_stats/data/blinded.csv") %>%
  filter(experiment == 1) %>%
  mutate(condition = ifelse(condition == "graph", "Graph", "No Graph"))

head(dataset)
```

We can visualise the data as such. We want to perform a comparision between the two groups of observed data and determine:

- whether they are different from each other
- if they are different, what is the magnitude of this difference

The Bayesian t-test allows us to answer these research questions. Like any other Bayesian data analysis, to implement this we need to perform three steps:

- first we need to describe the distribution of the observed data (also known as the Likelihood)
- next we need to define priors
- finally, we will compute the posteriors

We can then use the posteriors to compare between the two groups and answer the research questions. Let's us see how.

This is what the participants' responses look like:

```{r}
p1 = dataset %>%
  ggplot() +
  geom_point(aes(y = effectiveness, x = condition, color = condition), position = position_dodge2(width = 0.5), alpha = 0.2) +
  scale_colour_manual(values = c("#6405F9", "#FF7E50")) +
  coord_flip() +
  scale_y_continuous(breaks = seq(1, 9, by = 2), limits = c(0, 11)) +
  labs(y = "Effectiveness")

p2 = dataset %>%
  ggplot() +
  # geom_hline(yintercept = 0, size = 0.2, color = "#979797") +
  scale_y_continuous(breaks = 0, limits = seq(0, 1)) +
  theme(
    axis.title.x = element_blank(), 
    axis.line.x = element_blank(),
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.x = element_blank(),
    plot.margin = margin(0, .2, 0, .9, "cm")
  )

plot_grid(p2, p1, ncol = 1)
```
First, let's focus on the people in the "Graph" condition. We need to describe the observed data as a probability distribution. Since we are performing a t-test, we will use the Student's t distribution to describe the observed data.

```{r, warning = FALSE}
p1 = dataset %>%
  filter(condition == "Graph") %>%
  add_row(experiment = 1, condition = "No Graph", effectiveness = NULL) %>%
  ggplot() +
  geom_point(aes(y = effectiveness, x = condition), colour = "#6405F9", position = position_dodge2(width = 0.5), alpha = 0.2) +
  coord_flip() +
  scale_y_continuous(breaks = seq(1, 9, by = 2), limits = c(0, 11)) +
  labs(y = "Effectiveness")

p2 = dataset %>%
  ggplot() +
  geom_hline(yintercept = 0, size = 0.2, color = "#979797") +
  scale_y_continuous(breaks = 0, limits = seq(0, 1)) +
  theme(
    axis.title.x = element_blank(), 
    axis.line.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    plot.margin = margin(0, .2, 0, .9, "cm")
  )

plot_grid(p2, p1, ncol = 1)
```

```{r}
df.graph = filter(dataset, condition == "Graph")
N = nrow(df.graph)
data_mu = mean(df.graph$effectiveness)
data_sigma = sd(df.graph$effectiveness)
likelihood = tibble(x = seq(0, 11, by = 0.02), y = gamlss.dist::dTF2(x, data_mu, data_sigma/sqrt(N)))

p2 = ggplot() +
  geom_area(data = likelihood, aes(x, y), colour = "#0119F5", fill = "#0119F5", alpha = 0.1, size = 0.5) +
  annotate("text", label = "Likelihood =\nP(data | mean)", x = 8, y = 2) +
  geom_hline(yintercept = 0, size = 0.2, color = "#979797") +
  scale_y_continuous(breaks = 0, limits = c(0, 3)) +
  scale_x_continuous(breaks = seq(0, 9, by = 2), limits = c(0, 11)) +
  theme(
    axis.title.x = element_blank(), 
    axis.line.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    plot.margin = margin(0, .2, 0, .9, "cm")
  )

plot_grid(p2, p1, ncol = 1)
```

```{r, warning = FALSE}
prior_mu = 5
prior_sigma = 0.5
prior = tibble(x = seq(0, 11, by = 0.02), y = gamlss.dist::dTF2(x, prior_mu, prior_sigma))

p2 = ggplot() +
  geom_area(data = prior, aes(x, y), colour = "#75D6FF", fill = "#75D6FF", alpha = 0.3, size = 0.5) +
  geom_area(data = likelihood, aes(x, y), colour = "#0119F5", fill = "#0119F5", alpha = 0.1, size = 0.5) +
  annotate("text", label = "Likelihood =\nP(data | mean)", x = 8, y = 1) +
  annotate("text", label = "Prior =\nP(mean)", x = 3, y = 0.6) +
  geom_hline(yintercept = 0, size = 0.2, color = "#979797") +
  scale_x_continuous(breaks = seq(0, 9, by = 2), limits = c(0, 11)) +
  scale_y_continuous(breaks = 0, limits = c(0, 3)) +
  theme(
    axis.title.x = element_blank(), 
    axis.line.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    plot.margin = margin(0, .2, 0, .9, "cm")
  )

plot_grid(p2, p1, ncol = 1)
```

```{r}
get_post_mu = function (data, prior_mu, prior_sigma) {
  data_mu = mean(data)
  data_sigma = sd(data)
  n = length(data)
  ((prior_mu/prior_sigma^2) + ((n * data_mu)/data_sigma^2))/((1/prior_sigma^2) + (n/data_sigma^2))
}

get_post_sigma = function (data, prior_mu, prior_sigma) {
  data_sigma = sd(data)
  n = length(data)
  sqrt(1/((1/prior_sigma^2) + (n/data_sigma^2)))
}
```


```{r, warning = FALSE}
post_mu = get_post_mu(df.graph$effectiveness, prior_mu, prior_sigma)
post_sigma = get_post_sigma(df.graph$effectiveness, prior_mu, prior_sigma)
posterior = tibble(x = seq(0, 11, by = 0.01), y = gamlss.dist::dTF2(x, post_mu, post_sigma))

p2 = ggplot() +
  geom_area(data = prior, aes(x, y), colour = "#75D6FF", fill = "#75D6FF", alpha = 0.1, size = 0.5) +
  geom_area(data = likelihood, aes(x, y), colour = "#0119F5", fill = "#0119F5", alpha = 0.1, size = 0.5) +
  geom_area(data = posterior, aes(x, y), colour = "#6405F9", fill = "#6405F9", alpha = 0.3, size = 0.5) +
  annotate("text", label = "Likelihood =\nP(data | mean)", x = 8, y = 1) +
  annotate("text", label = "Prior =\nP(mean)", x = 3, y = 0.6) +
  annotate("text", label = "Posterior =\nP(mean | data)", x = 7.8, y = 2) +
  geom_hline(yintercept = 0, size = 0.2, color = "#979797") +
  scale_x_continuous(breaks = seq(0, 9, by = 2), limits = c(0, 11)) +
  scale_y_continuous(breaks = 0, limits = c(0, 3)) +
  theme(
    axis.title.x = element_blank(), 
    axis.line.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    plot.margin = margin(0, .2, 0, .9, "cm")
  )

plot_grid(p2, p1, ncol = 1)
```


```{r}
posterior1 = posterior

df.nograph = filter(dataset, condition == "No Graph")
post_mu2 = get_post_mu(df.nograph$effectiveness, prior_mu, prior_sigma)
post_sigma2 = get_post_sigma(df.nograph$effectiveness, prior_mu, prior_sigma)
posterior2 = tibble(x = seq(0, 11, by = 0.01), y = gamlss.dist::dTF2(x, post_mu2, post_sigma2))

p1 = dataset %>%
  ggplot() +
  geom_point(aes(y = effectiveness, x = condition, color = condition), position = position_dodge2(width = 0.5), alpha = 0.2) +
  scale_colour_manual(values = c("#6405F9", "#FF7E50")) +
  coord_flip() +
  scale_y_continuous(breaks = seq(1, 9, by = 2), limits = c(0, 11)) +
  labs(y = "Effectiveness")

p2 = ggplot() +
  geom_area(data = posterior1, aes(x, y), colour = "#6405F9", fill = "#6405F9", alpha = 0.2, size = 0.5) +
  geom_area(data = posterior2, aes(x, y), colour = "#FF7E50", fill = "#FF7E50", alpha = 0.2, size = 0.5) +
  geom_hline(yintercept = 0, size = 0.2, color = "#979797") +
  scale_x_continuous(breaks = seq(0, 9, by = 2), limits = c(0, 11)) +
  scale_y_continuous(breaks = 0, limits = c(0, 3)) +
  theme(
    axis.title.x = element_blank(), 
    axis.line.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    plot.margin = margin(0, .2, 0, .9, "cm")
  )

plot_grid(p2, p1, ncol = 1)
```


