---
title: "Introduction to Transparent Bayesian Data Analysis (chi2-course)"
date: "`r format(Sys.time(), '%d %B, %Y')`"
author: Abhraneel Sarma and Fumeng Yang
output: 
  html_document: 
    highlight: tango
    toc: yes
    toc_float: true
    toc_collapsed: true
    df_print: kable
---

```{=html}
<style>
.kable-table{
 overflow:scroll;
}

</style>
```


```{r knit-github, eval=FALSE,echo=FALSE,include=FALSE}
# compile for github, you can ignore this
library(rmarkdown)
render("bayesian-analysis.Rmd", md_document(variant = "gfm"), 
       knit_root_dir = "/Users/fm/Documents/Github/CHI-Course-Transparent-Quant")
```

```{r message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
rm(list = ls(all.names = TRUE)) 

library(dplyr)
library(tibble)
library(purrr)
library(ggpubr)
library(tidyr)
library(forcats)
library(gtools)

library(broom)
library(broom.mixed)

library(modelr)

library(brms)
library(tidybayes)

library(ggdist)
library(bayesplot)
library(ggplot2)
library(knitr)

BRM_BACKEND <- ifelse(require("cmdstanr"), 'cmdstanr', 'rstan')
```

```{r include=FALSE}
# specify color theme for the two poses
theme_green = '#20a598'
theme_blue = '#4574bf'
theme_yellow = '#fbc63d'
theme_red = '#e2524b'

scale_color_theme = \(...) scale_color_manual(..., 
  values = c('prior' = theme_yellow,
             'expansive' = theme_blue, 
             'posterior' = theme_green,
             'constrictive' = '#888888'),
  guide = "none",
  aesthetics = c("color", "fill")
)


# set up the global theme
theme_set(theme_ggdist() + 
          theme(strip.background = element_blank(),
               plot.title = element_text(hjust = .5)))

# We also define some ggplot styles
theme_density = theme(
    axis.text.y = element_text(colour = "#ffffff"),
    axis.title.y = element_blank(),
    axis.line.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.text.x = element_text(colour = "#ffffff"),
    axis.title.x = element_blank(),
    axis.line.x = element_blank(),
    axis.ticks.x = element_blank())

```

# Introduction

In this document, we will outline the Bayesian analogs of the statistical analyses described in lecture 1 ([Github code](https://github.com/chatchavan/CHI-Course-Transparent-Quant-Lecture-1)).

## Note on MathJax

If you use RStudio's Visual editor mode. Check if you can see anything written in the following parenthesis: ($k$). If not, your RStudio has a bug. This problem will be fixed in future versions. Meanwhile, if some text seems to be missing in the file below, simply position your cursor around the missing area to reveal the underlying code.

# Dataset

The following dataset is from experiment 2 of "How Relevant are Incidental Power Poses for HCI?" [(Jansen & Hornbæk, 2018)](https://doi.org/10.1145/3173574.3173588). Study participants were asked to either make an expansive posture or a constrictive posture before performing a task. The experiment investigated whether posture could potentially have an effect on risk taking behavior.

First, we load the data.

```{r load-dataset}
pose_df = readr::read_csv("data/poses_data.csv", show_col_types = FALSE) %>%
  mutate(condition) %>%
  group_by(participant)
```

The data has been aggregated for each participant: - `condition = expansive` indicates expansive posture, and `condition = constrictive` indicates constrictive posture - The dependent variable is `change` which indicates the percentage change in risk-taking behavior. Thus, it is a continuous variable.

For the purposes of this demo, we are only concerned with these two variables. We can ignore the other variables for now.

```{r show-dataset}
head(pose_df %>% select(participant, condition, change))
```

# Intuition of Bayesian Statistics

The Bayesian t-test (BEST) assumes that the data in the two conditions arises from two separate t-distributions. In the following section, we will describe the process for one of the conditions in the experiment.

We will use the $Normal(\mu = 20, \sigma = 20)$ as the prior distribution.

First, we define some functions for manual calculation of the posterior normal distribution:

```{r}
sigma_post = function(sigma_prior, sigma, n = 1) {
  sqrt(1 / (1 / (sigma_prior^2) + n / (sigma^2)))
}

mu_post = function(mu_prior, sigma_prior, mu, sigma, n = 1) {
  tau = sigma_post(sigma_prior, sigma, n)
  (tau^2 / sigma_prior^2)*mu_prior + (n * tau^2 / sigma^2)*mu
}
```

```{r}
d.p2 = tibble(
  group = c("prior", "expansive", "constrictive", "posterior"), 
  mu = c(20, 32.82, 31.61, mu_post(20, 20, 32.82, 7.52)), 
  sd = c(20, 7.52, 7.06, sigma_post(20, 7.52))
) %>%
  mutate(
    cutoff_group = list(c(1:7)),
    cutoff = list(c(0, 15, 25, 30, 32.82, 40, 100))
  ) %>%
  unnest(c(cutoff_group, cutoff)) %>%
  mutate(
    cutoff = if_else(group == 'prior', 100, cutoff),
    x = map(cutoff, ~ seq(from = -40, ., by = 0.1)),
    y = pmap(list(x, mu, sd), ~ dnorm(..1, ..2, ..3))
  )
```

In the plot below, we show the raw data distribution for the two conditions:

```{r, fig.height = 4, fig.width = 8}
p1 = pose_df %>%
  ggplot() +
  geom_point(aes(x = change, y = condition, colour = condition), 
             position = position_jitter(height = 0.1), alpha = 0.7) +
  scale_color_theme() + 
  labs(y = "Condition") +
  theme(
    legend.position = "none", 
    axis.line.y = element_blank(), 
    axis.ticks.y = element_blank(),
    axis.title.y = element_blank()
  ) +
  scale_x_continuous(limits = c(-40, 100), breaks = seq(-40, 100, by = 20))

p2.blank = tibble(y = c("expansive", "constrictive"), x = 0) %>%
  ggplot(aes(x, y)) +
  scale_color_theme() + 
  theme_density

cowplot::plot_grid(p2.blank, p1, nrow = 2)
```

First, we define a function to help us plot different cutoff groups.

```{r plot-preliminaries}
plot_preliminary <- function(data, group_num, group_name){
  data %>%
  filter(cutoff_group == group_num & group %in% group_name) %>%
  unnest(c(x, y)) %>%
  ggplot(aes(x, y)) +
  #geom_line(aes(color = group), size = 1) +
  # density
  geom_area(aes(fill = group, color = group), position = "identity", size = 1, alpha = 0.3) +
  scale_x_continuous(limits = c(-40, 100)) +
  scale_y_continuous(limits = c(0, 0.1)) +
  scale_color_theme() +
  theme_density
}

```

```{r, fig.height = 4, fig.width = 8}
cowplot::plot_grid(
  plot_preliminary(d.p2, 0, NULL), 
  p1, nrow = 2)
```

Next, we plot the prior density:

```{r, fig.height = 4, fig.width = 8}
cowplot::plot_grid(
  plot_preliminary(d.p2, 1, 'prior'), 
  p1, nrow = 2)
```

Then we describe step by step, how the likelihood is computed:

```{r, fig.height = 4, fig.width = 8}
cowplot::plot_grid(
  plot_preliminary(d.p2, 1, 'expansive'), 
  p1, nrow = 2)
```

```{r, fig.height = 4, fig.width = 8}
cowplot::plot_grid(
  plot_preliminary(d.p2, 2, 'expansive'), 
  p1, nrow = 2)
```

```{r, fig.height = 4, fig.width = 8}
cowplot::plot_grid(
  plot_preliminary(d.p2, 3, 'expansive'), 
  p1, nrow = 2)
```

```{r, fig.height = 4, fig.width = 8}
cowplot::plot_grid(
  plot_preliminary(d.p2, 5, 'expansive'), 
  p1, nrow = 2)
```

```{r, fig.height = 4, fig.width = 8}
cowplot::plot_grid(
  plot_preliminary(d.p2, 6, 'expansive'), 
  p1, nrow = 2)
```

```{r, fig.height = 4, fig.width = 8}
cowplot::plot_grid(
  plot_preliminary(d.p2, 7, 'expansive'), 
  p1, nrow = 2)
```



We want to compute the posterior, which is the product of the prior and likelihood:

```{r, fig.height = 4, fig.width = 8}
cowplot::plot_grid(
  plot_preliminary(d.p2, 7, c('prior', 'expansive', 'posterior')), 
  p1, nrow = 2)
```


Of course, if you have `gganimate` and `magick` working on your computer, you can try the following code to get a gif! Here, we are not going to show the code.

```{r fig.height = 4, fig.width = 8, message=FALSE, warning=FALSE, echo=FALSE}

if(require("gganimate") & require("magick")){
  
  library(gganimate)
  library(magick)
  
  # create the top animation of distribution
  ani <-  d.p2 %>%
    filter(group != 'constrictive') %>% 
    # force frame index
    mutate(frame = if_else(
     group == 'expansive' | group == 'prior', as.integer(cutoff_group),
     as.integer(case_when(
       # only show posterior as the last frame
       group == 'posterior' ~ 8,
       TRUE ~ 0
     ))
    )) %>% 
    # force the last frame to have prior and expansive
    rbind(
       d.p2 %>% filter((group == 'expansive' | group == 'prior') & cutoff_group == 7) %>% 
         mutate(frame = 8)
    ) %>% 
    unnest(c(x, y)) %>%
    ggplot(aes(x, y)) +
    # plot them
    geom_area(aes(fill = group, color = group),  position = "identity", size = 1, alpha = 0.3) +
    scale_x_continuous(limits = c(-40, 100)) +
    scale_y_continuous(limits = c(0, 0.1)) +
    scale_color_theme() +
    theme_density +
    transition_manual(frame)
  
  # render the gif
  ani_gif <- animate(ani, width = 800, height = 200)
  
  a_mgif <- image_read(ani_gif)
  
  # create the bottom part
  set.seed(1)
  bottom_ani <- pose_df %>%
    # some hack to force gganimate to create a gif
    mutate(frame = 1) %>%
  ggplot() +
  geom_point(aes(x = change, y = condition, colour = condition), 
             position = position_jitter(height = 0.1, seed = 1234), 
             alpha = 0.7, size = 2) +
  scale_color_theme() + 
  labs(y = "Condition") +
  theme(
    legend.position = "none", 
    axis.line.y = element_blank(), 
    axis.ticks.y = element_blank(),
    axis.title.y = element_blank()
  ) +
  scale_x_continuous(limits = c(-40, 100), breaks = seq(-40, 100, by = 20)) +
  transition_manual(frame)

  bottom_gif <- animate(bottom_ani, width = 800, height = 200)
  
  b_mgif <- image_read(bottom_gif)
  
  # combine them in each frame
  new_gif <- image_append(c(a_mgif[1], b_mgif[1]), stack = TRUE)
  for(i in 2:8){
    combined <- image_append(c(a_mgif[i], b_mgif[1]), stack = TRUE)
    new_gif <- c(new_gif, combined)
  }
  
  new_gif
}
```

# Model 1: Equal standard deviations

## Step 1: model specification

Fumeng: I hope to keep this example simple. Is there a __simple__ way to justify the choice of $\nu$? 



$$
\begin{align}
y_{i} &\sim \mathrm{Student\_t}(\mu, \sigma_{0}, \nu_{0}) \\
\mu &= \beta_{0} + \beta_{1} * x_i \\
\sigma_{0}  &\sim \mathrm{Normal}(0, 10) \\
\beta_{0} &\sim \mathrm{?} \\
\beta_{1} &\sim \mathrm{Normal}(0,2) \\
\nu_{0} &\sim \mathrm{?} \\
i & \in \{\mathrm{expnansive}, \mathrm{constrictive}\}
\end{align}
$$


### prior checks


## Step 2: model fitting


## Step 3: check posteriors


# Model 2: The BEST test model

## Step 1: model specification

This model is the BEST test model as described by Kruschke in the paper *Bayesian estimation supersedes the t-test*. In this model, $\beta$ indicates the mean difference in the outcome variable between the two groups (in this case, the percent change in the BART scores). We fit different priors on $\beta$ and set different weights on these priors to obtain our posterior estimate.

$$
\begin{align}
y_{i} &\sim \mathrm{T}(\nu, \mu, \sigma) \\
\mu &= \alpha_{0} + \beta * x_i \\
\sigma &= \sigma_{a} + \sigma_{b}*x_i \\
\beta &\sim \mathrm{Normal}(\mu_{0}, \sigma_{0}) \\
\sigma_a, \sigma_b &\sim \mathrm{Cauchy}(0, 2) \\
\nu &\sim \mathrm{exp}(1/30)
\end{align}
$$

# Model 3: Alternative SAYA model

SAYA = SArma + YAng

## model specification

$$
\begin{align}
y_{i} &\sim \mathrm{T}(\nu, \mu, \sigma) \\
\mu &= \alpha_{0} + \beta * x_i \\
\sigma &= \sigma_{a} + \sigma_{b}*x_i \\
\beta &\sim \mathrm{Normal}(\mu_{0}, \sigma_{0}) \\
\sigma_a, \sigma_b &\sim \mathrm{HalfNormal}(0, 10) \\
\nu &\sim \mathrm{exp}(1/30)
\end{align}
$$
