---
title: "Session 4 - Bayesian modeling (chi2023-course)"
date: "`r format(Sys.time(), '%d %B, %Y')`"
author: Fumeng Yang and Abhraneel Sarma
output: 
  html_document: 
    css: "CHI23course-style.css"
    toc_depth: 2
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: no
    df_print: kable
---


```{r setup, include=FALSE}
rm(list = ls(all.names = TRUE))
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, fig.width = 10, fig.height = 6)
```

# Introduction

# Load packages

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(brms)
library(tidybayes)
library(cowplot)
BRM_BACKEND <- ifelse(require("cmdstanr"), 'cmdstanr', 'rstan')

theme_set(theme_ggdist())
```




```{r}
df_between <- read_csv("df_between.csv")
```

```{r}
df_between
```


```{r}
ggplot(df_between) +
  stat_histinterval(aes(x = wpm, y = cond))
```


#  model 1

## Step 1
Use the get_prior function to check out the default priors:

```{r}
get_prior(wpm ~ cond,  data = df_between)
```


## Step 2

Update to proper priors 

```{r}
priors_choice <- c(
  prior(normal(0, 5), class = 'b'),
  prior(normal(50, 10), class = 'Intercept'),
  prior(normal(0, 2), class = 'sigma', lb = 0)
)
```


this is a placeholder for visualizing the priors


```{r}
m_prior_check <- brm(
  wpm ~ cond, 
  data = df_between,
  family = student(), 
  prior = priors_choice,
  seed = 1234,
  sample_prior = 'only'
)

```


Prior check

What they should see here
1. positive-only
2. on average, it should be around 50
3. symmetric
4. < 100

```{r}
newdata <-   crossing(cond = df_between$cond)
add_predicted_draws(
    newdata, 
    m_prior_check
) %>% 
ggplot(aes(
  x = .prediction, # put the correct variable here
  y = cond # 
)) +
stat_halfeye()
  
```



```{r}
m <- brm(
  wpm ~ cond,
  data = df_between,
  family = student(), 
  prior = priors_choice,
  seed = 1234
)

m <- brm(
  # write out the model formula 
  data = df_between,
  family = # specify family, 
  prior = priors_choice,
  seed = 1234
)

```

## Step 3 

```{r}
summary(m)
```

prediction?

```{r}

newdata %>%
  add_predicted_draws(m, ndraws = NULL, seed = 1234) %>% 
  ggplot() +
  stat_halfeye(
    aes(x = .prediction, y = cond), # should be changed
    position = position_nudge(y = .1),
    height = .8
  ) +
  geom_point(df_between,
             mapping = aes(x = wpm, y = cond),
             size = 5, shape = 20, alpha = .4) 
  
  
```

mean?

```{r}

posterior_draws <- 
newdata %>%
  add_epred_draws(m, seed = 1234)  
  
posterior_draws %>% 
  ggplot() +
  stat_halfeye(
    aes(
      x = .epred, 
      y = cond
    ),
    position = position_nudge(y = .1),
    height = .8
  ) +
  geom_point(df_between,
             mapping = aes(x = wpm, y = cond),
             size = 5, shape = 20, alpha = .4) 
  
  
```


```{r}
posterior_draws %>% 
  group_by(cond) %>% 
  median_qi(.epred)
```

```{r}
posterior_draws_diff <-
posterior_draws %>% 
  compare_levels(variable = .epred, by = cond, comparison = list(c('adaptive', 'constant'), c('off', 'constant'), c('adaptive', '')))
```



```{r}
posterior_draws_diff %>% 
  group_by(cond) %>% 
  median_qi()
```


```{r}
posterior_draws_diff %>% 
  ggplot() + 
  stat_halfeye(aes(x = .epred, y = cond)) + 
  geom_vline(xintercept = 0, linetype = 2)
```


# model 2

```{r}
df_between_more <- read_csv('df_between_more.csv')
```


## Step 1
Use the get_prior function to check out the default priors:

```{r}
get_prior(wpm ~ cond + (1|user_id),  data = df_between_more)
```


## Step 2

Update to proper priors 

```{r}
priors_choice <- c(
  prior(normal(0, 5), class = 'b'),
  prior(normal(50, 10), class = 'Intercept'),
  prior(normal(0, 2), class = 'sd', lb = 0),
  prior(normal(0, 2), class = 'sigma', lb = 0)
)
```


this is a placeholder for visualizing the priors


```{r}
m_rand_intercepts <- brm(
  wpm ~ cond + (1|user_id), 
  data = df_between_more,
  family = student(), 
  prior = priors_choice,
  seed = 1234
)

```


```{r}
m_rand_intercepts %>% 
  add_epred_draws(
    newdata,
    re_formula = NA # remove random intercepts
  ) %>% 
ggplot() +
  stat_slab(aes(x = .epred, y = cond))
```






