---
title: "Session 4 - Bayesian modeling (chi2023-course)"
date: "`r format(Sys.time(), '%d %B, %Y')`"
author: Fumeng Yang and Abhraneel Sarma
output: 
  html_document: 
    css: "CHI23course-style.css"
    toc_depth: 2
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: no
    df_print: kable
---


```{r setup, include=FALSE}
rm(list = ls(all.names = TRUE))
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, fig.width = 10, fig.height = 6)
```

# Introduction

# Load packages

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(brms)
library(rlang)
library(tidybayes)
library(cowplot)
BRM_BACKEND <- ifelse(require("cmdstanr"), 'cmdstanr', 'rstan')

theme_set(theme_ggdist())
```


# plot

```{r}
df_between <- read_csv("df_between.csv")
head(df_between)
```


```{r}
ggplot(df_between) +
  stat_histinterval(aes(x = wpm, y = cond))
```


### Step 1:
Use the get_prior function to check out the default priors:

```{r}
get_prior(
  # TODO: write down the model syntax ...
  ,  
  data = df_between
)
```


### Step 2:

Update to proper priors to 

```{r}
priors_choice <-  c(
    prior(
          # TODO: plug-in your choice here
          # hint: normal, student_t
          class = 'b'),
    prior(
          # TODO: plug-in your choice
          # hint: normal, student_t
          class = 'sigma', lb = 0)
  )
```

```{r}
priors_choice <- c(
  prior(normal(0, 5), class = 'b'),
  prior(student_t(3, 50, 10), class = 'Intercept'),
  prior(normal(0, 2), class = 'sigma', lb = 0)
)
```


this is a placeholder for visualizing the priors that abhraneel is working on...

```{r}
get_prior_dist = function(priors) {
  as_tibble(priors) |>
    mutate(prior_dist = map(prior, ~ eval_tidy(parse_expr(paste0("dist_", .)))))
}

get_prior_dist(priors_choice) |> 
  ggplot() +
  stat_dist_halfeye(aes(y = class, xdist = prior_dist))
```


## Prior check

```{r}
m_prior_check <- brm(
  # TODO: write down the model syntax ...
  data = df_between,
  family = # TODO: write down the likelihood function, 
  prior = priors_choice,
  seed = 1234,
  sample_prior = 'only'
)

```


What they should see here
1. positive-only
2. on average, it should be around 50
3. symmetric
4. < 100

```{r}
newdata <-   crossing(cond = df_between$cond)
add_predicted_draws(
    newdata, 
    m_prior_check
) %>% 
ggplot(aes(
  x = # TODO: enter the variable for predictions here
  y = # TODO: enter the variable for conditions here
)) +
stat_halfeye()
```

## Step 2: Running the model

```{r}
m <- brm(
  # TODO: write out the model formula 
  data = df_between,
  family =  # TODO: write down the likelihood function, 
  prior = priors_choice,
  seed = 1234
)
```

## Step 3: Checking the model results and reporting them

```{r}
summary(m)
```

## Posterior predictive checks

The code below implements a posterior predictive checks. We have provided the code which will extract posterior predictive draws from your model. You'll need to inspect the model, identify the variables and enter them to implement a posterior predictive check. 

```{r}
posterior_predictive_draws = newdata %>%
  add_predicted_draws(m, ndraws = NULL, seed = 1234)

posterior_predictive_draws %>% 
  ggplot() +
  stat_halfeye(
    aes(
      x = # TODO: enter the variable here
      y = # TODO: enter the variable here
    )
    position = position_nudge(y = .1),
    height = .8
  ) +
  geom_point(df_between,
             mapping = aes(x = # TODO: plug in x, 
                           y = # TODO: plug in y
                           ),
             size = 5, shape = 20, alpha = .4) 
  
  
```
## Posterior means for each of the three conditions


```{r}

posterior_draws <- 
newdata %>%
  add_epred_draws(m, seed = 1234)  
  
posterior_draws %>% 
  ggplot() +
  stat_halfeye(
    aes(
      x = # TODO: enter the variable here, 
      y = # TODO: enter the variable here
    ),
    position = position_nudge(y = .1),
    height = .8
  ) +
  geom_point(df_between,
             mapping = aes(x = # TODO: plug in x, 
                           y = # TODO: plug in y
                           ),
             size = 5, shape = 20, alpha = .4) 
  
  
```


```{r}
posterior_draws %>% 
  group_by( 
          # TODO: plug in the variable, 
         ) %>% 
  median_qi(
          # TODO: plug in the variable
  )
```

```{r}
posterior_draws_diff <-
posterior_draws %>% 
  compare_levels(variable = .epred, 
                 by = cond, 
                 comparison = list(c('adaptive', 'constant'), 
                                   c('off', 'constant'), 
                                   c('adaptive', 'off')))
```



```{r}
posterior_draws_diff %>% 
  group_by( 
           # TODO: plug in the variable, 
           ) %>% 
  median_qi(
           # TODO: plug in the variable
  )
```


```{r}
posterior_draws_diff %>% 
  ggplot() + 
  stat_halfeye(aes(
    x =  # TODO: plug in x, 
    y =  # TODO: plug in y, 
    )) + 
  geom_vline(xintercept = 0, linetype = 2)
```







